Predicting solutions for the Military Flight and Maintenance Planning problem

### 2020-09-04 R/presentations/EDAA202003/presentation.Rmd
changes to EDAA202003 ppt.

---
title: "An alternative MIP formulation for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "February, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

@@

<div class="my-footer"><span>ISAE-SUPAERO, Université de Toulouse, France    
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
ROADEF2020</span></div>

@@

1. Problem.
1. State of the art.
1. MIP formulations.
2. Machine Learning.
2. Experiments.
3. Conclusions.

@@

@@

* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the end of the horizon.

@@

### Previous formulation

* $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
* $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
* $u_{it}$    :  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$.

\begin{align}
 & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
 & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
 & rft_{it} \leq rft_{i(t-1)} - u_{it} + H^M m_{it}
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
& rft_{it} \in [0,H^M]
      & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:first-mu}
\end{align}

### New formulation

* $a_{ijtt'}$ : =1 if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
* $m_{ip}$ : =1 if aircraft $i$ uses check pattern $p$, zero otherwise.
  * each pattern $p$ has a single feasible combination of check starts for an aircraft during the whole planning (usually only 1-2 checks per aircraft).

\begin{align}
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{M} + M (1 - m_{ip}) & \notag \\
  & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
\end{align}

```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')

### Formulation

@@

### Is this better?

1. It uses 3 times the number of constraints and 3 times the number of variables.
  * variables: 11000 => 28000.
  * constraints: 13000 => 48000.
1. It is still better. Better lineal relaxation, better performance.

### Distance between maintenances

.pull-left[
* The distance between maintenance has a maximum of $E^{M}$ periods.
* Depending on the instance, the optimal distance can be shorter.
* This distance conditions the total number of patterns to create.
]

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

@@

1. Train a statistical model to predict the mean distance between maintenances for any given instance.
2. Use this information to limit all possible combinations of patterns to generate.

@@

1. **Performance**: a smaller model is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

@@

**The better we're able to predict the optimal distance between maintenances for the whole fleet, the less optimality we will lose**

@@

.pull-left[
* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Training**: 5000 small instances.
* **Input features**: 
  * mean flight demand per period,
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: mean distance between maintenances.
]

@@

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

>>>>

---
title: "Predicting solutions for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "September, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

@@

<div class="my-footer"><span>ISAE-SUPAERO, Université de Toulouse, France    
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
EDAA Journée doctorants 2020</span></div>

@@

1. Problem.
2. State of the art.
1. Solving techniques.
2. Machine Learning.
2. Experiments.
3. Conclusions.

@@

### Military aircraft maintenance

.pull-left[

* Mirage 2000 fleet.
* A single workshop to do the maintenance operations.
* Check type D (overhaul maintenances): 
  * every 5 years or 1000 flight hours and 
  * takes 4-6 months.
* Fleets of 15-60 aircraft.
* Planning horizons of 90 periods (= 7.5 years).

]

.pull-right[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('Mirage_2000C.jpg')
```

*French Mirage 2000C. Public Domain*
]

@@

* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the each period of the horizon.

@@

### "Classical" solving techniques

$a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
$m_{it_1t_2}$: aircraft $i$ starts a check in $t_1$ and another check in $t_2$.

@@

@@

1. Train a statistical model to predict characteristics of good (optimal?) solutions for any given instance.
2. Use this information to limit the solution space and make it easier to find those potentially good solutions.

@@

1. **Performance**: a smaller problem is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

@@

**The better we're able to predict those characteristics for the whole fleet, the less optimality we will lose**

@@

* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Training**: 5000 small instances.
* **Input features**: 
  * mean flight demand per period,
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: mean distance between maintenances.

@@

```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')
```
